PyTracer: Automatically Profiling Numerical Instabilities in Python

### 2022-05-16 IEEEtran/paper.tex
More edits

% \ifANNOTATED
\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\DeclareRobustCommand{\add}[1]{\textcolor{ao(english)}{#1}}%\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{\textcolor{lightred}{\sout{#1}}}
\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}
\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
% \else
% \DeclareRobustCommand{\add}[1]{#1}
% \DeclareRobustCommand{\remove}[1]{}
% \DeclareRobustCommand{\addCite}[1]{#1}
% \newcommand{\TG}[1]{}
% \fi

@@

@@

\remove{Listings~\mbox{\ref{fig:wrapper_creation}} and~\mbox{\ref{fig:generic_wrapper}} show
    \mbox{\pytracer}'s function instrumentation mechanism.}
For each function to instrument, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
that have a \texttt{\_\_call\_\_} attribute and are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning.

@@

        \pytracer trace parsing and aggregation, \emph{result}: precision for
        main outputs in significant bits. \valid: successful run, \remove{\warn: error
            from invalid noise injection,} \cross: numerical error or divergent
        traces. \textit{nan} stands for (\texttt{Not-A-Number}). Example names

@@

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on SciPy
and scikit-learn examples. Columns \textit{trace} and \textit{parse} report on
the tracing and aggregation steps with two possible outcomes: 1) green for
success, meaning that the execution ended well and no issues were raised by the
application, and 2) red for numerical error in the application, meaning that the
execution raised a Python error before the end. \remove{, and 3)
    orange for errors due to the Fuzzy noise injection that are
    not considered numerical errors, meaning that the execu-
    tion raised a Python error but due to a perturbation that
    should not have been introduced (i.e. a floating-point value
    representing an integer).}
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples\remove{ and showed an average
    precision loss of 8 bits}. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. 
\remove{The instrumentation with Full MCA is
    more invasive and reduced the number of successful exe-
    cutions to 12/40. Among the 28/40 failed executions, one
    failed due to an actual numerical error (Bayesian Ridge
    Regression example) and the other ones failed due to
    perturbations that should not have been introduced.}

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
the precision for the main outputs of SciPy and scikit-learn examples. We
measured the precision in significant bits on the final primary example output
by taking the element-wise mean for non-scalar values and, if the example
produces many outputs, the average value among them. The numerical
quality across all examples varied from 0 to 52, with an average of 44
significant bits. We note that only one example failed due to a
numerical error (\texttt{face\_recognition}
\remove{for RR and \texttt{Bayesian Ridge Regression} for full MCA}).

@@

\add{PyTracer supports native floating point formats as well as compound data types,
    such as complex floating point values, N-dimensional arrays and arbitrary Python
    classes.}
Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} \add{show the
    visualization of the \texttt{fft\_1D} input and output values. Results are
    complex values encoded in double precision. PyTracer allows visualizing the
    numerical precision of real (blue points) and imaginary (orange points) parts.}
\add{The} top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs}
show the input mean value while the bottom row shows the standard deviation. The
points with low magnitude in Figure~\ref{fig:fft1D_inputs} correspond to inputs
near 0 when the input of $\sin$ or $\exp$ is close to $k\pi$, $k \in
    \mathbb{Z}$. Indeed, since MCA introduces a slight noise, the result is not
exactly 0. We can see in Figure~\ref{fig:fft1D_inputs} that the maximal noise
introduced by MCA on the inputs is in the order of $10^{-14}$, two orders of
magnitude higher than the $ulp \simeq 10^-16$ for double precision. This slight
noise introduced can be interpreted as white noise on the input signal. We can
see on the bottom row of Figure~\ref{fig:fft1D_outputs} that the frequency noise
is of the same order of of magnitude as the input noise, which is expected. The
two peaks at x=38 and x=562 corresponds to the actual frequencies of the input
signal.

Figure~\ref{fig:spline2d_rr} \add{shows the results of the spline evaluation with
    \texttt{bisplev} from the \texttt{spline\_2D} example. It computes the 2D spline
    interpolation for function $z=(x+y)e^{-6(x^2+y^2)}$, with $(x,y) \in
        [-1,1]\times[-1,1]$. The result is a 2D-array that is natively supported by
    PyTracer. We observe a significant loss of precision in some regions, which follows
    an interesting pattern that could be further investigated.}

@@

Replacing \texttt{gesdd} for \texttt{gesvd} solved the SVD convergence issue.
\pytracer allows to quickly visualize the new precision of the result.
Figure~\ref{fig:face_recognition_svd} shows the precision of the eigenvalues
computed displayed in decreasing order with \texttt{gesvd} \add{in single precision}, 
which varies between 15 and 20
significant bits out of the 24 available in single precision. The classifier
metrics (precision, recall, F1-score, and support) are identical to the IEEE
execution for all MCA samples. \add{We can note that even with the \texttt{gesvd} method,
    the precision of the singular values and eigenvectors is not uniform. 
    Indeed, we can see that the precision of the lowest singular values and eigenvectors decreases 
    until reaching 0 significant bits.}

The second example compares a Bayesian Ridge Regression (BRR) to the Ordinary
Least Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
non-significant. \pytracer traces reveal that the SVD solver used is once again the root
cause for this error. \add{Indeed, the BayesianRidge model first computes an SVD
    before iteratively updating the coefficients until reaching the convergence
    criterion. However, PyTracer shows an issue in the singular values computed.}
\add{We can see in Figure}~\ref{fig:BRR_svd} \add{that the last singular values is
    \texttt{NaN} due to numerical errors in the SVD, although the remaining
    singular values exhibit high numerical precision (between 44 and 50 bits).
    Surprinsingly, this \texttt{NaN} value does not raise an exception and is silently
    propagated. Indeed, the stopping criterion is never verified and the 
    optimization stops after reaching the maximal number of iterations.
    Figure}~\ref{fig:BRR_svd_fixed} \add{shows that uses
    the \texttt{gesvd} instead of the \texttt{gesdd} method removes the \texttt{NaN} value
    although the last singular values stills have a low precision (below 0).
    Following Equation}~\ref{eq:sig-digits}, \add{a number of significant digits below 0 means that the
    variability ($\sigma$) is higher than the average value ($\mu$)}.

\add{Similarly, for OLS, PyTracer tracers show that the
    regression coefficients fitted are \texttt{NaN} values. By looking at the callgraph, we can
    see that the scipy calls the \texttt{lstsq} method from the
    \texttt{scipy.linalg} package which uses the \texttt{gesdd} LAPACK function.
    Figure}~\ref{fig:OLS_svd} \add{shows that one singular value is a \texttt{NaN} value,
    which corrupted the calculation of the coefficients. When using the \texttt{gesvd} method, 
    the \texttt{NaN} value disapears 
    (Fig.}~\ref{fig:OLS_svd_fixed}\add{), resulting in a precision of
    42 bits in average in the coefficients. We can see that having
    access to all the values computed by the application with PyTracer allows
    the user to efficiently identify numerical issues.}

In conclusion, \pytracer and the MCA model were helpful to detect the
instability of the \texttt{gesdd} method mentioned in several GitHub issues,
including the LAPACK repository. Furthermore, it is worth mentioning that a Pull
Request has recently been submitted to let the user select the SVD method and
avoid this kind of issue. Thus, \pytracer can detect instabilities that affect
fundamental scientific computing components such as the LAPACK library. \add{The
    unified view of all traces allows the user to quickly navigate through the
    numerical precision of different variables.}
% The Pytracer findings can therefore be extended to many application domains. %
% \tristan{conclusion needs to be discussed. Note to self: generaliser un peu la
% conclusion: pytracer permet de se rendre compte de problemes de stabilite dont
% plusieurs librairies sont dependantes.}

\begin{figure*}
    \centering
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_U_sig.pdf}
        \caption{Left matrix $U$ of the SVD decomposition}
        \label{fig:randomized_svd_U}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/svd_ret_S_sig.pdf}
        \caption{Singular values $\Sigma$ of the SVD decomposition}
        \label{fig:randomized_svd_S}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_V_sig.pdf}
        \caption{Right matrix $V$ of the SVD decomposition}
        \label{fig:randomized_svd_V}
    \end{subfigure}
    \caption{Precision of the \texttt{randomized\_svd} function in the
        \texttt{face\_recognition} example using the \texttt{gesvd} SVD method
        and RR. Figures~\ref{fig:randomized_svd_U},~\ref{fig:randomized_svd_S},
        and~\ref{fig:randomized_svd_V} show respectively the values $U$,
        $\Sigma$ and $V$. Numerical precision is lower for smaller singular
        values and this loss of precision translates to singular vectors. It
        would be interesting to investigate the use of numerical precision for
        dimensionality reduction. }
    \label{fig:face_recognition_svd}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values.png}
        \caption{BRR (\texttt{gesdd})}
        \label{fig:BRR_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values.png}
        \caption{OLS (\texttt{gesdd})}
        \label{fig:OLS_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values_fixed.png}
        \caption{BRR (\texttt{gesvd})}
        \label{fig:BRR_svd_fixed}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values_fixed.png}
        \caption{OLS (\texttt{gesvd})}
        \label{fig:OLS_svd_fixed}
    \end{subfigure}
    \caption{\add{Numerical precision of the singular values of the SVD from the
            Bayesian Ridge model fitting (left) and from the Ordinary Least Square model
            fitting (right) in the Bayesian Ridge Regression example. Red stars
            represent \texttt{NaN} values. 
            Switching from Divide \& Conquer (\texttt{gesdd}) to QR (\texttt{gesvd}) factorization improves numerical precision.
            PyTracer helps visualizing improvements.}}
\end{figure}

@@

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} \remove{and \textit{Slowdown MCA}} shows the slowdown within the Fuzzy environment.
Without Fuzzy, \pytracer introduces an average slowdown of $17\times$ for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, the \remove{Fuzzy instrumentation} \add{using MCA
    operations (through Fuzzy) instead of the native ones} increases the average
slowdown to $37\times$ for RR\remove{and $43\times$ for Full MCA}.

@@

Finally, the wide range of results precision observed in our experiments
suggests that such
analyses would benefit from being conducted systematically. We showed that
well-known scientific libraries are prone to numerical errors even though they
have been well tested for decades. The prevalence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

@@

In the future, we plan to extend the post-processing analysis to aggregate
traces when the execution flow path diverges by leveraging techniques used in
code coverage analysis. Moreover, we would like to extend \pytracer to
instrument elementary arithmetic operations. \remove{Finally, although \pytracer supports multiple noise models,
we only tested it with Monte-Carlo arithmetic. Therefore, we would like to
compare different models against MCA.} \pytracer is available at
\href{https://github.com/big-data-lab-team/pytracer}{github://big-data-lab-team/pytracer}
under MIT License.

>>>>

% \ifANNOTATED
\definecolor{ao(english)}{rgb}{0.0, 0.5, 0.0}
\DeclareRobustCommand{\add}[1]{\textcolor{ao(english)}{#1}}%\sethlcolor{lightgreen}\hl{#1}}
\DeclareRobustCommand{\remove}[1]{}
\DeclareRobustCommand{\addCite}[1]{\textcolor{lightgreen}{\mbox{#1}}}
\newcommand{\TG}[1]{\color{red}\textsc{From Tristan}: #1\xspace\color{black}}
% \else
% \DeclareRobustCommand{\add}[1]{#1}
% \DeclareRobustCommand{\remove}[1]{}
% \DeclareRobustCommand{\addCite}[1]{#1}
% \newcommand{\TG}[1]{}
% \fi

@@

\add{
The remainder of this section assumes some knowledge of Python concepts
such as classes, functions, modules, decorators, callables, and iterators.
We refer to~\cite{beazley2013python} for a detailed presentation of these
concepts.
}

@@

\remove{Listings~\mbox{\ref{fig:wrapper_creation}} and~\mbox{\ref{fig:generic_wrapper}} show
    \mbox{\pytracer}'s function instrumentation mechanism.}
For each function to instrument, \pytracer's \texttt{loader} class creates a
wrapper function, dynamically compiles it with the \texttt{compile} builtin
function, and substitutes it for the original module function.  Although simple,
this technique does not apply to callable class instances, i.e., class instances
\add{that can be called directly with the () operator, because they,} are not of type
\texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify
the type of the class instance to \texttt{FunctionType}, which could cause
syntax and semantic errors, and 2) mask class attributes, leading to
\texttt{AttributeError} exceptions. \pytracer overrides the
\texttt{\_\_call\_\_} attribute with the wrapper function when possible to
overcome this issue. When the \texttt{\_\_call\_\_} attribute is not writable,
Pytracer does not instrument the class and returns a warning.

@@

        \pytracer trace parsing and aggregation, \emph{result}: \add{result} precision
         in number of significant bits. \valid: successful run, \remove{\warn: error
            from invalid noise injection,} \cross: numerical error \add{(tracing)} or 
            \add{different control flow paths (parsing).} \textit{nan} stands for (\texttt{Not-A-Number}). Example names

@@

Table~\ref{tab:pytracer_results_summary} summarizes \pytracer's outcome on
SciPy and scikit-learn examples. Columns \textit{trace} and \textit{parse}
report on the tracing and aggregation steps with two possible outcomes:
\add{green \valid\xspace for success and red \cross\xspace for error. Tracing errors 
occur in the application as a result of 
numerical perturbations, whereas parsing errors occur in \pytracer when traces result from different 
control flow paths. 
} \remove{, and 3)
    orange for errors due to the Fuzzy noise injection that are
    not considered numerical errors, meaning that the execu-
    tion raised a Python error but due to a perturbation that
    should not have been introduced (i.e. a floating-point value
    representing an integer).}
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random
Rounding mode was entirely successful for 32/40 examples\remove{ and showed an average
    precision loss of 8 bits}. The remaining examples failed due to numerical errors
in the libraries, which we discuss hereafter. 
\remove{The instrumentation with Full MCA is
    more invasive and reduced the number of successful exe-
    cutions to 12/40. Among the 28/40 failed executions, one
    failed due to an actual numerical error (Bayesian Ridge
    Regression example) and the other ones failed due to
    perturbations that should not have been introduced.}

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes
the \add{result} precision for the main outputs of SciPy and scikit-learn examples. We
measured the \add{result} precision in significant bits on the final primary example output
by taking the element-wise mean for non-scalar values and, if the example
produces many outputs, the average value among them. \add{Significant bits are computed on the final program output 
and can therefore be computed even when parsing fails.} \add{We use the term ``result precision"
to denote a statistical variability measured in number of significant bits, in 
contrast with the other meanings of the word precision used in expressions
``virtual precision" or ``single/double precision'' in this paper.} The numerical
quality across all examples varied from 0 to 52, with an average of 44
significant bits. We note that only one example failed due to a
numerical error (\texttt{face\_recognition}
\remove{for RR and \texttt{Bayesian Ridge Regression} for full MCA}).

@@

\add{PyTracer supports native floating point formats as well as compound data types,
    such as complex floating point values, N-dimensional arrays and arbitrary Python
    classes.}
Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs} \add{show the
    visualization of the \texttt{fft\_1D} input and output values. Results are
    complex values encoded in double precision. PyTracer allows visualizing the
    \add{result} precision of real (blue points) and imaginary (orange points) parts.}
\add{The} top row in Figures~\ref{fig:fft1D_inputs} and~\ref{fig:fft1D_outputs}
show the input mean value while the bottom row shows the standard deviation. \add{Even though we 
used Random Rounding (RR), function inputs are perturbed since they result from multiple intermediate 
computations perturbed with RR.} The
points with low magnitude in Figure~\ref{fig:fft1D_inputs} correspond to inputs
near 0 when the input of $\sin$ or $\exp$ is close to $k\pi$, $k \in
    \mathbb{Z}$. Indeed, since MCA introduces noise, the result is not
exactly 0. We can see in Figure~\ref{fig:fft1D_inputs} that the maximal noise
introduced by MCA on the inputs is in the order of $10^{-14}$, two orders of
magnitude higher than the $ulp \simeq 10^-16$ for double precision. This slight
noise introduced can be interpreted as white noise on the input signal. We can
see on the bottom row of Figure~\ref{fig:fft1D_outputs} that the frequency noise
is of the same order of of magnitude as the input noise, which is expected. The
two peaks at x=38 and x=562 corresponds to the actual frequencies of the input
signal.

Figure~\ref{fig:spline2d_rr} \add{shows the results of the spline evaluation with
    \texttt{bisplev} from the \texttt{spline\_2D} example. It computes the 2D spline
    interpolation for function $z=(x+y)e^{-6(x^2+y^2)}$, with $(x,y) \in
        [-1,1]\times[-1,1]$. The result is a 2D-array that is natively supported by
    PyTracer. We observe a significant loss of \add{result} precision in some regions, which follows
    an interesting pattern that could be further investigated.}

@@

Replacing \texttt{gesdd} for \texttt{gesvd} solved the SVD convergence issue.
\pytracer allows to quickly visualize the new \add{result} precision.
Figure~\ref{fig:face_recognition_svd} shows the \add{result} precision of the eigenvalues
computed with \texttt{gesvd} \add{in single precision}. \add{The result precision}
varies between 15 and 20
significant bits out of the 24 available in single precision. The classifier
metrics (precision, recall, F1-score, and support) are identical to the IEEE
execution for all MCA samples. \add{We can note that even with the \texttt{gesvd} method,
    the \add{result} precision of the singular values and eigenvectors is not uniform. 
    Indeed, we can see that the \add{result} precision of the lowest singular values and eigenvectors decreases 
    until reaching 0 significant bits.}

The second example compares a Bayesian Ridge Regression (BRR) to the Ordinary
Least Squares (OLS) estimator on a synthetic dataset and for one-dimensional
regression using polynomial feature expansion. Although the example does not
raise a runtime error, the regression coefficients computed from the fitting are
non-significant. \pytracer traces reveal that the SVD solver used is once again the root
cause for this error. \add{Indeed, the BayesianRidge model first computes an SVD
    before iteratively updating the coefficients until reaching the convergence
    criterion. However, PyTracer shows an issue in the singular values computed.}
\add{We can see in Figure}~\ref{fig:BRR_svd} \add{that the last singular values is
    \texttt{NaN} due to numerical errors in the SVD, although the remaining
    singular values exhibit high \add{result} precision (between 44 and 50 bits).
    Surprinsingly, this \texttt{NaN} value does not raise an exception and is silently
    propagated. Indeed, the stopping criterion is never verified and the 
    optimization stops after reaching the maximal number of iterations.
    Figure}~\ref{fig:BRR_svd_fixed} \add{shows that uses
    the \texttt{gesvd} instead of the \texttt{gesdd} method removes the \texttt{NaN} value
    although the last singular values stills have a low \add{result} precision (below 0).
    Following Equation}~\ref{eq:sig-digits}, \add{a number of significant digits below 0 means that the
    variability ($\sigma$) is higher than the average value ($\mu$)}.

\add{Similarly, for OLS, PyTracer tracers show that the
    regression coefficients fitted are \texttt{NaN} values. By looking at the callgraph, we can
    see that the scipy calls the \texttt{lstsq} method from the
    \texttt{scipy.linalg} package which uses the \texttt{gesdd} LAPACK function.
    Figure}~\ref{fig:OLS_svd} \add{shows that one singular value is a \texttt{NaN} value,
    which corrupted the calculation of the coefficients. When using the \texttt{gesvd} method, 
    the \texttt{NaN} value disapears 
    (Fig.}~\ref{fig:OLS_svd_fixed}\add{), resulting in a \add{result} precision of
    42 bits in average in the coefficients. We can see that having
    access to all the values computed by the application with PyTracer allows
    the user to efficiently identify numerical issues.}

In conclusion, \pytracer and the MCA model were helpful to detect the
instability of the \texttt{gesdd} method mentioned in several GitHub issues,
including the LAPACK repository. Furthermore, it is worth mentioning that a Pull
Request has recently been submitted to let the user select the SVD method and
avoid this kind of issue. Thus, \pytracer can detect instabilities that affect
fundamental scientific computing components such as the LAPACK library. \add{The
    unified view of all traces allows the user to quickly navigate through the
    \add{result} precision of different variables.}
% The Pytracer findings can therefore be extended to many application domains. %
% \tristan{conclusion needs to be discussed. Note to self: generaliser un peu la
% conclusion: pytracer permet de se rendre compte de problemes de stabilite dont
% plusieurs librairies sont dependantes.}

\begin{figure*}
    \centering
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_U_sig.pdf}
        \caption{Left matrix $U$ of the SVD decomposition}
        \label{fig:randomized_svd_U}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/svd_ret_S_sig.pdf}
        \caption{Singular values $\Sigma$ of the SVD decomposition}
        \label{fig:randomized_svd_S}
    \end{subfigure}
    \begin{subfigure}{0.3\linewidth}
        \includegraphics[width=\linewidth]{figure/face_recognition/randomized_svd_ret_V_sig.pdf}
        \caption{Right matrix $V$ of the SVD decomposition}
        \label{fig:randomized_svd_V}
    \end{subfigure}
    \caption{Precision of the \texttt{randomized\_svd} function in the
        \texttt{face\_recognition} example using the \texttt{gesvd} SVD method
        and RR. Figures~\ref{fig:randomized_svd_U},~\ref{fig:randomized_svd_S},
        and~\ref{fig:randomized_svd_V} show respectively the values $U$,
        $\Sigma$ and $V$. \add{Result} precision is lower for smaller singular
        values and this loss of precision translates to singular vectors. It
        would be interesting to investigate the use of numerical precision for
        dimensionality reduction. }
    \label{fig:face_recognition_svd}
\end{figure*}

\begin{figure}
    \centering
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values.png}
        \caption{BRR (\texttt{gesdd})}
        \label{fig:BRR_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values.png}
        \caption{OLS (\texttt{gesdd})}
        \label{fig:OLS_svd}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/BRR_singular_values_fixed.png}
        \caption{BRR (\texttt{gesvd})}
        \label{fig:BRR_svd_fixed}
    \end{subfigure}
    \begin{subfigure}{0.49\linewidth}
        \includegraphics[width=\linewidth]{figure/OLS_singular_values_fixed.png}
        \caption{OLS (\texttt{gesvd})}
        \label{fig:OLS_svd_fixed}
    \end{subfigure}
    \caption{\add{Result} precision of the singular values of the SVD from the
            Bayesian Ridge model fitting (left) and from the Ordinary Least Square model
            fitting (right) in the Bayesian Ridge Regression example. Red stars
            represent \texttt{NaN} values. 
            Switching from Divide \& Conquer (\texttt{gesdd}) to QR (\texttt{gesvd}) factorization improves numerical precision.
            PyTracer helps visualizing improvements.}
\end{figure}

@@

\pytracer's tracing overhead consists of (1) module instrumentation time
(initialization overhead) and (2) data write time during the execution (runtime
overhead). The initialization overhead is proportional to the number of
functions and classes traced. For example, we measured an average initialization
time of 8 seconds during our experiments with  NumPy, SciPy, and Scikit-learn.
This overhead can be reduced by excluding intensively used functions such as
\texttt{numpy.array} from the instrumentation.
Figure~\ref{fig:performance_tracing} summarizes \pytracer's tracing overhead for
SciPy and scikit-learn examples. \textit{Slowdown} represents the slowdown of
\pytracer without fuzzy measured as the ratio between the actual execution time
(without \pytracer) and the execution time using \pytracer. The \textit{Slowdown
    RR} \remove{and \textit{Slowdown MCA}} shows the slowdown within the Fuzzy environment.
Without Fuzzy, \pytracer introduces an average slowdown of $17\times$ for
the examples running less than 8 seconds because \pytracer spends most of its
time in the initialization step. Since this step is a constant factor of the
number of functions instrumented, we also measured the instrumentation cost
itself (\textit{Slowdown (Instrumentation)}) calculated by subtracting the
initialization time to \textit{Slowdown}. We then obtain an average slowdown of
$3.5\times$. Finally, \remove{the Fuzzy instrumentation} \add{using MCA
    operations (through Fuzzy) instead of the native ones} increased the average
slowdown to $37\times$\remove{for RR and $43\times$ for Full MCA}.

@@

Finally, the wide range of result precision observed in our experiments
suggests that such
analyses would benefit from being conducted systematically. We showed that
well-known scientific libraries are prone to numerical errors even though they
have been well tested for decades. The prevalence of these libraries in
scientific codes should thus impel users to integrate numerical analysis in
their development workflow in addition to testing their implementation.

@@

In the future, we plan to extend the post-processing analysis to aggregate
traces \add{obtained with different control flow paths,} by leveraging techniques used in
code coverage analysis. Moreover, we would like to extend \pytracer to
instrument elementary arithmetic operations. \remove{Finally, although \pytracer supports multiple noise models,
we only tested it with Monte-Carlo arithmetic. Therefore, we would like to
compare different models against MCA.} \pytracer is available at
\href{https://github.com/big-data-lab-team/pytracer}{github://big-data-lab-team/pytracer}
under MIT License.
